---
title: "Programming Assessment 1 - Reproducible Research"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

## Loading and preprocessing the data


```{r, echo=TRUE}

act <- read.csv("C:\\Users\\Mark\\Downloads\\repdata_data_activity\\activity.csv")

library(dplyr)

a <- filter(act,!is.na(steps))
b <- group_by(a,date)
c <- summarise(b,totalsteps = sum(steps))
c <- as.data.frame(c)
c$date <- as.Date(c$date)
```

## Histogram of the total number of steps taken each day


```{r histogram_1, echo=TRUE}
library(ggplot2)

ggplot(c,aes(x=totalsteps)) + geom_histogram(col="black",fill="blue",alpha=0.2) + labs(title="Histogram for total days for step ranges", x="Steps", y="Days")
```

## Mean and median number of steps taken each day


```{r, echo=TRUE}
mean_median <- summarise(c,mean = mean(totalsteps),median = median(totalsteps))
print(mean_median)
```

## Time series plot of the average number of steps taken


```{r time_series_1, echo=TRUE}
d <- group_by(a,interval)
e <- summarise(d,averagesteps = mean(steps))
with(plot(interval,averagesteps,type="l",main="Avg Steps per interval",ylab="Average steps"),data=e)
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo=TRUE}
print(as.integer(e[which.max(e$averagesteps),1]))
```

## Code to describe and show a strategy for imputing missing data

```{r, echo=TRUE}
act_2 <- inner_join(act,e,by="interval")
act_2$steps[is.na(act_2$steps)] <- act_2$averagesteps
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r histogram_2, echo=TRUE}
f <- group_by(act_2,date)
g <- summarise(f,totalsteps = sum(steps))
g <- as.data.frame(g)
g$date <- as.Date(g$date)

ggplot(g,aes(x=totalsteps)) + geom_histogram(col="black",fill="blue",alpha=0.2) + labs(title="Histogram for total days for step ranges accounting for NAs", x="Steps", y="Days")
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r panel_1, echo=TRUE}
act_2$date <- as.Date(act_2$date)
act_2 <- mutate(act_2,weekday = ifelse(weekdays(act_2$date) == "Sunday" | weekdays(act_2$date) == "Saturday","weekend","weekday"))

h <- group_by(act_2,interval,weekday)
i <- summarise(h,newaveragesteps = mean(steps))

library(lattice)
xyplot(newaveragesteps ~ interval | weekday,i,type='l',main="Average Steps per Interval",ylab="Average steps")
```
